---
title: "Life Tables using HistData"
author: "Alex Delamer"
format: 
    html:
        theme: flatly

execute: 
  echo: false
  message: false
  warning: false
---

```{r setup}
library(HistData)
library(tidyverse)
library(kableExtra)
```

### Filtering Breslau data

Filtering for observations where age is greater than or equal to 5.

```{r filtering}
#| echo: true

breslau_filtered <- Breslau |> 
    filter(age >= 5)
```

#### Summary of filtered data

The below table (@tbl-breslau) shows the average age of people who died in the city of Breslau (now Wroclaw) in Poland between the years 1687 and 1691 inclusive.

```{r summary}
#| label: tbl-breslau
#| tbl-cap: Average age of death in Breslau, 1687-1691

breslau_summary <- breslau_filtered |> 
    pivot_longer(cols = starts_with("year"), names_to = "year", values_to = "count") |> 
    mutate(year = str_remove(year, "year")) |> 
    group_by(year) |> 
    summarise(average_age = round(weighted.mean(age, count), 2), total_deaths = sum(count))

kable(breslau_summary) |> 
    kable_styling(bootstrap_options = c("striped", "hover"), fixed_thead = TRUE)
```

#### Plot of mortality in Breslau

As we can see in @fig-breslau, the average age of death for citizens of Breslau between 1687 and 1691 stayed relatively consistent.

```{r}
#| label: fig-breslau
#| fig-cap: Average age of death in Breslau, 1687-1691

ggplot(breslau_summary, aes(x = year, y = average_age)) +
    geom_col(fill = "steelblue") +
    labs(x = "Year", y = "Average age of death")
```